---
title: "Exploring gapminder & dplyr"
author: "Ni Jie"
date: "2/8/2021"
output: 
  github_document: default
  html_document: default
---

<!-- BMW: See above for how to specify options for multiple output formats. -->

```{r setup, include = FALSE}
# BMW: Name the first chunk setup and put all library() calls in it.
knitr::opts_chunk$set(echo = TRUE) # BMW: Do not set error = TRUE
library(gapminder) # BMW:
library(palmerpenguins)
library(ggplot2)
library(tidyverse) # BMW: library(tidyverse) also loads dplyr and ggplot2, so you don't need both
library(dplyr) # BMW: Don't include View() calls or similar in your final script 
```

# Loading Packages

```{r load packages}
library(gapminder)
library(palmerpenguins)
library(ggplot2)
library(tidyverse)
library(dplyr)
View(gapminder)
View(palmerpenguins::penguins)
```

# Exercise 1: Basic dplyr

1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r}
subset1970 <- filter(gapminder, country %in% c("Austria", "Benin", "Brazil") & year >=1970 & year < 1980)
subset1970
```

<!-- BMW: in markdown, you need a blank line between paragraphs. A single line break is ignored. -->

1.2 Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r}
subset1970 %>%
  select(country, gdpPercap)
```

1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy.

```{r}
gapminder %>%
  mutate(changeLifeExp = lifeExp - lag(lifeExp)) %>%

  filter(changeLifeExp < 0)
```

1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country.

```{r}
gapminder %>%
  group_by(country) %>%

  filter(gdpPercap == max(gdpPercap)) # BMW: The question asked you to _filter_ the dataset, keeping the other variables, not just compute the max for each country.
```

1.5 Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

```{r}
gapminder %>%
  filter(country == "Canada") %>%

  ggplot(aes(x = lifeExp, y = log(gdpPercap))) + 
  geom_point()
```

# Exercise 2: Explore two variables with dplyr and ggplot2

2.1 I use "palmerpenguins::penguins" dataset and two quantitative variables that are "bill length" and "flipper length".

```{r}
#descriptive table

palmerpenguins::penguins %>% # BMW: The select() was unnecessary. Don't select unnecessarily. 
  summarize(mean_bill = mean(bill_length_mm, na.rm = TRUE),
            mean_flipper = mean(flipper_length_mm, na.rm = TRUE),
            sd_bill = sd(bill_length_mm, na.rm = TRUE),
            sd_flipper = sd(flipper_length_mm, na.rm = TRUE))

# scatterplot
palmerpenguins::penguins %>%

  ggplot(aes(x=bill_length_mm, y=flipper_length_mm)) + 
  geom_point(alpha = .5)

```

2.2 I pick "body mass" and "species". Histogram and Boxplot are used to visualize these variables.

```{r}
#descriptive table
palmerpenguins::penguins %>%
  group_by(species) %>%
  summarize(sample_n = n(), 
            mean_bodymass = mean(body_mass_g, na.rm = TRUE), 
            sd_bodymass = sd(body_mass_g, na.rm = TRUE))

#histogram
palmerpenguins::penguins %>%

  ggplot(aes(x = body_mass_g, fill = species)) + # BMW: you probably wanted fill
  geom_histogram(bins = 10, color = "black") +
  facet_wrap(~ species)


#boxplot
palmerpenguins::penguins %>%
  select(species, body_mass_g) %>%
  ggplot(aes(x=species, y=body_mass_g)) + 
  geom_jitter(aes(color = species), height = 0, width = .2) + # BMW: Always show the actual data, not only a boxplot or bar chart
  geom_boxplot(fill = "transparent") 
```

