---
title: "Consequences of Covid-19 for Candle Critiques"
author: "Ni Jie"
date: "3/31/2021"
output: html_document:
  code_folding: hide
  highlight: haddock
  theme: readable
---

```{r setup, include=FALSE}
library(yaml)
library(rmarkdown)
library(knitr)
library(dplyr)
library(tidyverse)
library(readxl)
library(gridExtra)
```


## Dataset Introduction

```{r}
scented <- read_xlsx(here::here("hw05", "Scented_all.xlsx"))
unscented <- read_xlsx(here::here("hw05", "Unscented_all.xlsx"))
```

This report presents the data analysis about impact of Covid-19 on scented candle ratings. The dataset contains the review and rating record of the Top 3 selling scented and unscented candles sold on Amazon. Below is a brief codebook of the relevant information: 

  1. `CandleID`: A numeric id variable for candle 1-3 in each category.
  2. `Date`: The date of the review.
  3. `Rating`: The 1-5 rating for the review.
  4. `Review`: For the scented candles, the text of the review.


## Part 1: Average daily ratings for scented and unscented candles

```{r}
#Step 1: Summarize data
daily_s <-
  scented %>%
  filter(Date > as.Date("2017-01-01")) %>%
  group_by(Date) %>%
  mutate(Average_Daily_Rating = mean(Rating)) %>% 
  na.omit() %>%
  select(CandleID, Date, Average_Daily_Rating)
daily_un <-
unscented %>%
  filter(Date > as.Date("2017-01-01")) %>%
  group_by(Date) %>%
  mutate(Average_Daily_Rating = mean(Rating)) %>% 
  na.omit() %>%
  select(CandleID, Date, Average_Daily_Rating)
#Step 2: Plotting
plot_daily_s <-
  daily_s %>%
  ggplot(
    aes(x = as.Date(Date), y = Average_Daily_Rating, 
        fill = as.factor(CandleID), color = as.factor(CandleID))) +
      geom_point(alpha = .25) +
      scale_color_manual(values = c("#F0E442", "#0072B2", "#D55E00")) + 
      scale_x_date(date_breaks = "6 months", date_labels = "%m-%Y") +
      geom_smooth(method = "loess") +
      geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dotted") + 
      geom_text(aes(x=as.Date("2020-01-20"), label = "2020-01-20", y = 1)) + 
      theme_minimal() + 
      labs(x = "Date", y = "Average Daily Rating") + 
      ggtitle("Top 3 Selling Scented Candles' Daily Ratings from 2017 to 2020")
plot_daily_un <- 
  daily_un %>%
  ggplot(
    aes(x = as.Date(Date), y = Average_Daily_Rating, 
        fill = as.factor(CandleID), color = as.factor(CandleID))) +
      geom_point(alpha = .25) +
      scale_color_manual(values = c("#F0E442", "#0072B2", "#D55E00")) + 
      scale_x_date(date_breaks = "6 months", date_labels = "%m-%Y") +
      geom_smooth(method = "loess") +
      geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dotted") +
      geom_text(aes(x=as.Date("2020-01-20"), label = "2020-01-20", y = 1)) + 
      theme_minimal() + 
      labs(x = "Date", y = "Average Daily Rating") + 
      ggtitle("Top 3 Selling Unscented Candles' Daily Ratings from 2017 to 2020")
grid.arrange(plot_daily_s, plot_daily_un, ncol = 1)
```

By plotting the average daily rating of both scented and unscented candles from 2017 to 2020 in 6-month increment, 
we can find the effect of Covid-19's outbreak in the US. Before 2020-01-20, the ratings of top 3 selling scented candles are 
consistently around 4 out of 5 points. After the Covid outbreak, more people lost their sense of smell which correlate to the 
unusual decreasing in the scented candles' rating.


## Part 2: Combined plot

```{r combine datasets, eval=FALSE}
candles_full <- bind_rows(
	scented = daily_s, 
	unscented = daily_un, 
	.id = "Scented"
)
```

```{r}
plot_full <-
  candles_full %>%
  ggplot(
    aes(x = as.Date(Date), y = Average_Daily_Rating, 
        fill = as.factor(Scented), color = as.factor(Scented), shape = as.factor(Scented))) +
      geom_point(alpha = .25) +
      scale_color_manual(values = c("#F0E442", "#0072B2")) + 
      scale_x_date(date_breaks = "6 months", date_labels = "%m-%Y") +
      geom_smooth(method = "loess", aes(linetype = as.factor(Scented))) + 
      geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dotted") +
      geom_text(aes(x=as.Date("2020-01-20"), label = "2020-01-20", y = 1)) + 
      theme_minimal() + 
      labs(x = "Date", y = "Average Daily Rating") + 
      ggtitle("Top 3 Selling Unscented vs. Scented Candles' Daily Ratings from 2017 to 2020")
plot_full
```

I believe the combined plot is much easier to interpret because the research interest is on the scented and unscented 
candles not the three brands of them. By looking at the second plot, both regression lines are on the same x and y scales, 
which makes the comparison easier. In general, we get the same conclusion from the combined plot as the separate plots.


## Part 3: Proportion of reviews mentioning "no scent"

```{r no scent function, include=FALSE}
str_no_scent <- function(x){
  as.numeric(
  	case_when(
  		str_detect(x, "[Nn]o scent") ~ "1", 
  		str_detect(x, "[Nn]o smell") ~ "1",
  		str_detect(x, "[Dd]oes not smell like") ~ "1",
  		str_detect(x, "[Dd]oesn't smell like") ~ "1",
  		str_detect(x, "[Cc]an't smell") ~ "1",
  		str_detect(x, "[Cc]annot smell") ~ "1",
  		str_detect(x, "[Ff]aint smell") ~ "1",
  		str_detect(x, "[Ff]aint scent") ~ "1",
  		str_detect(x, "[Dd]on't smell") ~ "1",
  		str_detect(x, "[Ll]ike nothing") ~ "1",
  		TRUE ~ "0"
  	)
  )
}
```

```{r}
#Step 1: Summarize Data
data_review <- 
  scented %>% 
  filter(Date >= as.Date("2019-01-01")) %>%
  mutate(month = lubridate::floor_date(Date, "month")) %>%
  mutate(noscent = str_no_scent(Review)) %>%
  select(month, noscent)
#Step 2: Plotting
plot_review <-
  data_review %>%
  ggplot(
    aes(x = as.Date(month), y = noscent)) + 
      stat_summary(aes(y = noscent), fun.data = mean_cl_normal) + 
      scale_x_date(date_breaks = "6 months", date_labels = "%m-%Y") + 
      geom_smooth(method = "loess") + 
      geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), color = "#0072B2") +
      geom_text(aes(x=as.Date("2020-01-20"), label = "2020-01-20", y = 1)) + 
      theme_minimal() + 
      labs(x = "Month", y = "Proportion of No-Scent Reviews") + 
      ggtitle("Top 3 Selling Scented Candles' DMonthly Reviews from 2019 to 2020")
plot_review
```

There is an increasing in proportion of reviews that mentioned "no scent" in 2020. Back in 2019, there are a consistent number 
of reviews mentioning no scent.
