---
title: "Homework03"
author: "Ni Jie"
date: "2/18/2021"
output: 
  github_document: default
  html_document: default
---
```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = FALSE)
```

# Loading Packages

```{r load packages}
library(dplyr)
library(tidyr)
library(tibble)
library(naniar)
library(scales)
library(ggplot2)
library(gridExtra)
library(tidyverse)
```

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
View(dictionary)
View(green_data)
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
  
```{r}
summary(green_data)
#recode -99 as missing values & recode student
green_recode <-
  green_data %>%
  replace_with_na_all(~.x == -99) %>%
  mutate(
    student = recode(student, "1" = "nonstudent", "2" = "student"))
summary(green_recode)
View(green_recode)
```   
## According to the summary, some of the items have -99 as the minimum values so I recoded them as NA.

  2. Compute total scores for the four scales.
     Recode variables as needed.
  
```{r}
#pull reversed items
reversed <- dictionary %>%
  filter(Keying == -1 | Keying == -2) %>%
  pull(Item)
reversed
#reverse
green_reversed <-
  green_recode %>%
  mutate(
    across(all_of(reversed),
           ~ recode(.x, "5" = 1, "4" = 2, "3" = 3, "2" = 4, "1" = 5)))
View(green_reversed)
#calculate total scores
green_total<-
  green_reversed %>% 
  rowwise() %>% 
  mutate(green_tot = mean(c_across(green1:green5)),
         comp_tot = mean(c_across(comp1:comp10)),
         intel_tot = mean(c_across(intel1:intel10)),
         open_tot = mean(c_across(open1:open10))) %>% 
  na.omit() %>%
  select(id, student, green_tot, comp_tot, intel_tot, open_tot) 
  View(green_total)
  summary(green_total)
```   
  
  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
     
```{r}
green_rescale <-
  green_total %>%
  mutate(
    green_pomp = rescale(green_tot, to=c(0,100), from=c(1,5)),
    comp_pomp = rescale(comp_tot, to=c(0,100), from=c(1,5)),
    intel_pomp = rescale(intel_tot, to=c(0,100), from=c(1,5)),
    open_pomp = rescale(open_tot, to=c(0,100), from=c(1,5))) %>%
  select(id, student, green_pomp, comp_pomp, intel_pomp, open_pomp)
  View(green_rescale)
  summary(green_rescale)
```  
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  
```{r}
#histogram
hist_green <-
  ggplot(green_rescale, aes(x=green_pomp)) + 
  geom_histogram(bins = 25)
hist_comp <-
  ggplot(green_rescale, aes(x=comp_pomp)) + 
  geom_histogram(bins = 25)
hist_intel <-
  ggplot(green_rescale, aes(x=intel_pomp)) + 
  geom_histogram(bins = 25)
hist_open <-
  ggplot(green_rescale, aes(x=open_pomp)) + 
  geom_histogram(bins = 25)
grid.arrange(hist_green, hist_comp, hist_intel, hist_open)
```  
## The scores for these four variables are generally normally distributed. The distributions for comp_pomp and open_pomp are skewed to the right while there are some outliers in the distribution of intel_pomp.
  
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
     
```{r}
# scatterplot
scatter_comp <-
  ggplot(green_rescale, aes(x=comp_pomp, y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))
scatter_intel <-
  ggplot(green_rescale, aes(x=intel_pomp, y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))
scatter_open <-
  ggplot(green_rescale, aes(x=open_pomp, y=green_pomp, 
                            color=student)) + 
  theme_bw() +
  geom_point(alpha = .5) +
  geom_smooth(aes(group=student))
  grid.arrange(scatter_comp, scatter_intel, scatter_open)
```     
## The overall relationships between green reputation and personalities are shared with students and nonstudents. Higher the score on compassion, intellect, and openness, higher the green reputation. However, for students, intellect and openness have a stronger effect on the green reputation.

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  
```{r}
ggplot(green_rescale, aes(x = green_pomp, color=student), alpha = .5) +
  geom_bar() +
  geom_density() +
  geom_point(aes(y = "idetity", color=student), alpha = .5)
```     

## The overall density of green reputation for both student and nonstudent groups are similar. While non students has a wider distribution on the green reputation scale and has a larger amount of them with scores around the mean.

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}
green_rescale %>%
  select(student, green_pomp, comp_pomp, intel_pomp, open_pomp) %>%
  group_by(student) %>%
  summarize(
    across(c(green_pomp, comp_pomp, intel_pomp, open_pomp),
           list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE),
                med = ~median(.x, na.rm = TRUE),
                min = ~min(.x, na.rm = TRUE),
                max = ~max(.x, na.rm = TRUE))))
```     

## Nonstudents have a slightly higher green reputation mean and smaller standard variance, as well as higher median, a larger range. Nonstudents also show a lower score on compassion despite their minimum score of compassion is much higher than the students. Besides, Nonstudents' minimum score on all personalities are much higher than students'.
